---
title: "Predicting Churn"
author: "Jamel Thomas"
date: "8/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Although there are many examples, churn prediction is one of the classical applications of Data Science that works. Churn prediction gives businessmen and bussinesswomen the power to catch those consumers who are likely to leave the company. They can in turn take appropriate measure to keep their business. In this project, we consider a dataset of 7,043 observations. The goal here is to predict, with a fair amount of accuracy, which observations are likely to churn. 

```{r echo=F}
churn <- read.csv("/Users/Jamel/Downloads/WA_Fn-UseC_-Telco-Customer-Churn.csv", header = T)
```



```{r barplot, echo = F}
#Make a function for barplots

plot.barplot <- function (x, xname = NULL, fillname = NULL, prop = F){
  if(!("package:ggplot2" %in% search()) )require(ggplot2)
  if(!(is.null(xname) | is.null(fillname))){
    t <- table(x[[xname]], x[[fillname]])
    if(prop)   t <- prop.table(t)
    df <- data.table::data.table(t)
    colnames(df) <- c(paste(xname), paste(fillname), "Frequency")
    if( is.null( names(x) ) ){
     return("Please pass a data.table() with column names")
    }else{
      g <- ggplot(df, aes(x = df[[xname]], y = Frequency, fill = df[[fillname]])) + 
        geom_bar(stat = "identity", position = "dodge") + ylab("Frequency") + 
        xlab(paste(xname) )+ guides(fill=guide_legend(title=paste(fillname))) + 
        theme_minimal()
      print(g)
      return(t)
    }
  }else{
    return("Please use both a xname and fillname")
  }
}

```

## Exploratory Data Analysis


```{r echo = F, warning=F}
require(ggplot2)
require(magrittr) #Load Packages
library(flipPlots)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r }
DT::datatable(churn[,c(1:2,16:21)])

```



## Including Plots

You can also embed plots, for example:

```{r, fig.align="center"}


my.data <- cbind(churn,data.frame(num = 1) )
my.data$SeniorCitizen <- factor(my.data$SeniorCitizen)
my.data2 <- aggregate(num~. , data = my.data[, c(7:8, 12, 16,21, 22)], sum)
  
  

SankeyDiagram(my.data2[, - dim(my.data2)[2]],
              link.color = "Source", 
              weights = my.data2$num) 

temp <- churn[churn$Churn == "Yes", ]
var <- temp$Contract  # the categorical data 

nrows <- 10
df <- expand.grid(y = 1:nrows, x = 1:nrows)
categ_table <- round(table(var) * ((nrows*nrows)/(length(var))))
categ_table[1] <- categ_table[1] - 1 #Adjust

df$category <- factor(rep(names(categ_table), categ_table))  
# NOTE: if sum(categ_table) is not 100 (i.e. nrows^2), it will need adjustment to make the sum to 100.

## Plot
ggplot(df, aes(x = x, y = y, fill = category)) + 
  geom_tile(color = "black", size = 0.5) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), trans = 'reverse') +
  scale_fill_brewer(palette = "Set3") +
  labs(title="Customer churn % (+)", subtitle="Contract Type",
       caption="Telecom") + 
  theme(
    plot.title = element_text(size = rel(1.2)),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank(),
    legend.position = "right")



```


```{r, echo=F, fig.align='center'}
temp <- churn[churn$Churn == "No", ]
var <- temp$Contract  # the categorical data 

nrows <- 10
df <- expand.grid(y = 1:nrows, x = 1:nrows)
categ_table <- round(table(var) * ((nrows*nrows)/(length(var))))

df$category <- factor(rep(names(categ_table), categ_table))  
# NOTE: if sum(categ_table) is not 100 (i.e. nrows^2), it will need adjustment to make the sum to 100.

## Plot
ggplot(df, aes(x = x, y = y, fill = category)) + 
  geom_tile(color = "black", size = 0.5) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), trans = 'reverse') +
  scale_fill_brewer(palette = "Set3") +
  labs(title="Customer churn % (-)", subtitle="Contract Type",
       caption="Telecom") + 
  theme(
    plot.title = element_text(size = rel(1.2)),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank(),
    legend.position = "right")



```

### Exploring 

```{r, fig.align="center"}

plot.barplot(x = churn, 
             xname = "Churn",
             fillname = "Contract")
plot.barplot(churn, 
             xname = "Churn" ,
             fillname = "Dependents")
plot.barplot(churn, 
             xname = "Churn", 
             fillname = "StreamingMovies")
plot.barplot(churn, 
             xname = "Churn", 
             fillname = "PaymentMethod") 

ggplot(churn, aes(x = Churn, y = tenure, fill= Contract)) + 
  geom_boxplot() + theme_minimal()
ggplot(churn, aes(x = Churn, y = tenure)) + 
  geom_boxplot() + theme_minimal()
ggplot(churn, aes(x = Churn, y = tenure, fill= TechSupport)) + 
  geom_boxplot() + facet_wrap(~SeniorCitizen)+ theme_minimal()

ggplot(churn, aes(x = Churn, y = MonthlyCharges, fill= Contract)) + 
  geom_boxplot() + theme_minimal()

ggplot(churn, aes(x = tenure, y = TotalCharges, color = Contract)) + geom_point( ) + geom_smooth()+ theme_minimal()
ggplot(churn, aes(x = MonthlyCharges, y = TotalCharges, color = Contract)) + geom_point( ) + geom_smooth()+ theme_minimal()

ggplot(churn, aes(x = tenure, y = MonthlyCharges, color = Churn)) + 
  geom_point( )+ geom_smooth()+ facet_wrap(~factor(InternetService))+ theme_minimal()

ggplot(churn, aes(x = Churn, y = tenure, fill = InternetService)) + 
  geom_boxplot() + facet_wrap(~TechSupport, nrow = 1)+ theme_minimal()

ggplot(churn, aes(x = MonthlyCharges, fill = Churn)) + 
  geom_histogram() + theme_minimal()#Monthly Charges by Churn

ggplot(churn, aes(x = MonthlyCharges, fill = TechSupport)) + 
  geom_histogram() + theme_minimal()
ggplot(churn, aes(x = MonthlyCharges, fill = InternetService)) + 
  geom_histogram() + theme_minimal()
ggplot(churn, aes(x = TotalCharges, fill = Contract)) + 
  geom_histogram() + facet_wrap(~InternetService) + theme_minimal()
 ggplot(churn, aes(x = tenure, y = MonthlyCharges, color = InternetService)) + 
  geom_point( )+ geom_smooth()+ theme_minimal()

```

## Model Building

For this mini-project, we will use logistic regression, decision trees, and k-nearest neighbors.
```{r, fig.align="center"}
set.seed(100)
indx <- sample(1:nrow(churn), .8*nrow(churn), replace = F)

```

### Logistic Regression
Since I am only using logistic regression to explore relationships, and I'm not necessarily too concerned with the accuracy of this particular model, I will hold off on splitting the data 80-20 until the subsequent models. 

```{r}

churn$Churn <- factor(churn$Churn)
churn$SeniorCitizen <- factor(churn$SeniorCitizen )
fit <- glm(Churn ~ SeniorCitizen + MonthlyCharges + tenure + Contract +
             PaymentMethod +TechSupport, data = churn, family = "binomial")
summary(fit)
exp(coef(fit))
exp(cbind(OR = coef(fit), confint(fit)))


```
Some intersting things to point out: based on the data, those who pay with elecronic check have a 50% higher odds of churning over the baseline. Moreover, a one unit increase in montly charges cost corresponds with a 2% greater odds of churning.

```{r, fig.align="center"}

tree = rpart::rpart(Churn ~ SeniorCitizen + MonthlyCharges + tenure + Contract +
               PaymentMethod +TechSupport + Dependents , data = churn)
rpart.plot::rpart.plot(tree)



train <- churn[indx,c("MonthlyCharges", "tenure", "TotalCharges", "Churn")]
test <- churn[-indx,c("MonthlyCharges", "tenure", "TotalCharges", "Churn")]

train <- na.omit(train)
test <- na.omit(test)
kn <- class::knn(train[,-4], test[,-4],train[,4] , k = 22, l = 4, prob = F, use.all = TRUE)
length(kn)
dim(test)

df <- table(Prob = kn, Obs = test[,"Churn"]) %>% data.frame()

ggplot(df, aes(x = Prob, y = Freq, fill = Obs)) + 
  geom_bar(stat = 'identity', position = "dodge") + theme_light()

table(Prob = kn, Obs = test[,"Churn"])
table(Prob = kn, Obs = test[,"Churn"]) %>% prop.table()


```

This final model was able to achieve about 80% accuracy. KNN is very competitve with logistic regression and decision trees out of the box model.
