---
title: "Plotly for College Analysis"
author: "Jamel Thomas"
date: "04 August 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plotly for College Analysis

Community colleges for many is a stepping stone to a larger dream of four year university. However, many of the risk factors that prevent students from succeeding in school are highly prevelant in the student body of 2-year institutions. 

In this post, we explore open college data. This dataset includes the name of the college, the aggregated number of people who passed a particular subject by transfer level and race. 

Let's read our data and take a look:

There is a wonderful package called `DT` that creates interactive plots in HTML. It even comes with a search feature. Check it out. 
```{r dataload, warning=FALSE, echo=F, results='hide'}
url <- "https://www.jamelt.com/data/CA_REMEDIAL_2014-2016.txt"

```

```{r cars, warning=FALSE, echo=T, results='hide'}
#Load Packages:
require(DT)
require(plotly)

school <- read.table(url, header = T)

```


```{r dt, warning=F, echo=T, eval=T}
DT::datatable(school)
```


## ggplot
The ggplot package is essentially the bread and butter of beautiful and easy to use plots. We will start with `data = school` (duh), and `x = ethnicity, y = sum`. The "sum"" here is just the total number of students who passed given course. 

```{r pressure, echo=T, fig= T, fig.align='center', fig.width=10}
g <- ggplot(data = school,
            aes(x = ethnicity, y = sum, fill = group)) + geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) #Vertical Axis
g
```

## Interactivity
In order to make this plot interactive, we have to do something really challenging. Are you ready? Here we go...

Type the following line of code. 

```{r plotly, echo=T, fig= T, fig.align='center', fig.width=10}
ggplotly(g)
```

That's it! We can now do all sorts of amazing plots with minimal code! Let's look as some more stuff. 


```{r bar, echo=T, fig= T, fig.align='center', fig.width=10}
g <- ggplot(data = school,
            aes(x = ethnicity, y = success, fill = ethnicity)) +
  geom_bar(stat = "identity")  +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) #Vertical Axis
ggplotly(g)

```

Make sure you clean your data!

```{r scatter, echo=T, fig= T, fig.align='center', fig.width=10}
#Refactor
school$group <- factor(school$group,
                       levels = c("One Level Below Transfer",
                                  "Two Levels Below Transfer",
                                  "Three Levels Below Transfer",
                                  "Four Levels Below Transfer"))

g <- ggplot(data = school,
            aes(x = students, y = sum, color = ethnicity)) +
  geom_point() + facet_wrap( ~ group, ncol=2)
ggplotly(g)

```


```{r scatter2, echo=T, fig= T, fig.align='center', fig.width=10}

g <- ggplot(data = school,
            aes(x = subject, y = sum, color = ethnicity)) +
  geom_boxplot() + facet_wrap( ~ ethnicity, ncol=2)
ggplotly(g)

```


```{r scatter3, echo=T, fig= T, fig.align='center', fig.width=10}

g <- ggplot(data = school,
            aes(x = group, y = sum, fill = subject)) +
  geom_boxplot() + facet_wrap( ~ subject, ncol=2)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) #Vertical Axis
ggplotly(g)
```