---
title: "Python in R Markdown?"
author: "Jamel Thomas"
date: "8/5/2018"
output: html_document
---

## Python in R Markdown

R Markdown supports Python integration with the package `reticulate`.The nice thing about R Markdown is that it is the first scripting language I have ever learned. Therefore, I am almost too comfortable with it. It is my go to for most of my project because it's so easy to get a proof of concept with new ideas. The downfall of it - until now - is that I was unaware that we could use PYTHON IN R MARKDOWN! Mind Blown! It sounds so easy too; duh, you could use Python in R Markdown this is 2018. But honestly, it has never crossed my mind to try... until now.

It is fairly easy to get started. Just install the package and run! Here is a list of all supported languages.

```{r eval = T }
names(knitr::knit_engines$get())

```

```{r eval = F }
install.packages("reticulate")
library(reticulate)
```

You will need to use `"``` {python} CODE CHUNK HERE ```" ` to get started.

```{python}
import sys
print(sys.version)
```

## Including Plots

You can also embed plots, for example:

```{python}
import matplotlib.pyplot as plt
plt.plot([6,8,10, 12, 21, 44,47, 50, 51])
plt.show()
```

<!-- And there you have it! You can eve mix and match! If you have a global variable, `y`, we can manipulate it in R or Python. -->


<!-- `PYTHON` -->
<!-- ```{python} -->
<!-- y = 10 -->
<!-- print(y) -->
<!-- ``` -->









### Go and create!