---
title: "SQL, Databases & R"
author: "Jamel Thomas"
date: "8/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SQL: 

SQL is a querying language used to manipulate databases. On its own, it is a powerful tool in any Data Scientist's toolbox. When you pair SQL with R (or Python), you have indespensible foundations to conduct even the basic analytics. This is not so much a lesson on SQL, but a lesson on connecting SQL with R via various packages. Moreover, we will use the packages `dbplyr` and `dplyr` to manipulate the datasets in a later blog. 

### Ready..... FIGHT!

## Intro
When you are using small data, it is fairly easy to store it locally. However, this can get more and more cumbersome as your data grows.
Using databases instead of storing all of your data on a flat file in memory is a lot more efficient because you can query only the data you need at any given time. 

R has the ability to connect to almost any database type via pacakages. The widely used types can range from MySQL, PostgreSQL and vSQL. We will explore the `schools` database found on my website and github. 

```{r packages, eval=T, echo=FALSE}
#install.packages(c("RSQLite"))
url<- "https://jamelt.com/data/CA_REMEDIAL_2014-2016.txt"
```

https://jamelt.com/data/CA_REMEDIAL_2014-2016.txt

```{r pressure, echo=T}
library(sqldf)

school <- read.table(url, header = T)

```

Let's first take a look at the data:

```{r}
head(school)
str(school)
```

Now to use SQL, we just add the string query in the `sqldf()` function


```{r}
query <- "SELECT count(college) FROM school WHERE ethnicity = 'African American' GROUP BY subject"
sqldf(
  query
)


```

The `GROUP BY` here is an aggregator that will summarize the data by a selected factor of two or more levels. 

```{r}
library(ggplot2)
DF = sqldf(
  "SELECT * FROM school WHERE ethnicity = 'African American' "
)
qplot(DF$sum, data = DF, geom = 'histogram') + xlab("Passed Class")
```


