---
title: "Where have I been?"
author: "Jamel Thomas"
date: "8/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Where Am I from?

I began this blog in the summer of 2018, mostly because I needed a place to put all of the things I do for fun. What better way than to build your own webssite and put all that crap there?? I have come across this beautiful post from the D3js website. It creates a plot of where the author has been. I will recreate the same plot here, but with the leaflet package in R. 


```{r dataload, echo = F, eval=T, results="hide"}
url <- "https://jamelt.com/data/cities-lived.csv"
# url2 <- "https://jamelt.com/data/stateslived.csv"


```

We will load in the data for the cities and states that I've lived. 

```{r data}
require(leaflet) #packages
require(DT)
require(rgdal)

cities <- read.csv(url, header = T)
DT::datatable(cities)
```

Leaflet comes with an easy to use pipeline: `%>%`. For example, given two functions `as.character()` and `as.numeric()`, we could code this two different ways. 

```{r pipe}
x <- factor(c("12.99", "99.01", "20"))

#Method one:
as.numeric( as.character(x) )

#Method two:

x %>% as.character %>% as.numeric

#But incorrect output of factor to numeric
as.numeric(x)
```

This is important to note because we will utilize the pipe notation. 

We have already loaded leaflet and the datasets, we just need to use them!

We will first load the leaflet base plot and add tiles:
```{r leafletBase, fig.align='center'}
L <- leaflet(data = cities) %>% 
  addTiles()
L
```

When we add out points, the mapview is automatically updated as well!

```{r mapview, fig.align='center'}
L %>%
  addCircles(lat = ~lat, lng = ~lon, radius = ~years*1000 , stroke = F, fillOpacity = 1)
```




